---
title: "Script No. 3"
author: "Marlene Kammerer"
date: "18 Oktober 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = TRUE
)

```

# Network positions, roles, and characteristics {.tabset}

## Tutorial

### Content

Well known already!!

  - Load packages 
  - Load the network data
  - Create network objects
  - Add actor attributes
  - Plot networks using centrality and actor type (--> we use a new package)
  
NEW
  - Assortativity (Looking for homophily)
  - Finding Bridges
  - A package to identify structural holes

### Install & load packages

Again, we work with the`igraph` package. But also with several other packages ... if you don't have them yet installed, do so!


```{r}

# install.packages("migraph") # to install the package 
library(igraph) # to load the package
```

### Load the network data

We will work again with the Swiss dataset. Since this is all well-known already. We do this VERY quickly now. Attention! We work with one new dataset, the two-level dataset.

Hence, we have several datasets:

  - A dataset that contains information in the organization name and it's type (covariate)
  - A dataset that contains information on the beliefs, instrument preferences, and secondary aspects that an actor holds (also covariates)
  - A dataset that contains information on whether an actors is involved as well at the level of international climate politics
  
Have a look at these datasets and try to understand what all these variables mean!

```{r}

## Attributes

beliefs.Ch <- read.csv2("Data/CH_beliefs.csv", header=TRUE, fileEncoding = "latin1", stringsAsFactors = FALSE)
twolevel.ch <- read.csv2("Data/CH_2level.csv", header=TRUE, fileEncoding = "latin1", stringsAsFactors = FALSE)

## Collaboration network

coll.ch <- read.csv2("Data/cH_coll.csv", header=TRUE)
rownames(coll.ch) <- coll.ch$X
coll.ch$X <- NULL

## Influence network

inf.ch <- read.csv2("Data/cH_influence.csv", header=TRUE)
rownames(inf.ch) <- inf.ch$X
inf.ch$X <- NULL

## Actor type

types.ch <- read.csv2("Data/CH_type.csv", header = TRUE, fileEncoding = "latin1", stringsAsFactors = FALSE)

```

### Create network objects

As usual we have to create network objects! This time, we only work with the collaboration network. We directly add the type variable and the organization name to the network.


```{r}

coll.ch.mat <- as.matrix(coll.ch) # transform dataset into matrix format
coll.ch.net <- graph.adjacency(coll.ch.mat, mode="directed", diag=FALSE) # transform matrix into graph object
V(coll.ch.net)$type <- types.ch$Type
V(coll.ch.net)$klar <- types.ch$Klarname_EN
list.vertex.attributes(coll.ch.net)

```
### Plot network

We plot the collaboration network. This time using a new package. To do so, we need to transform the the igraph object into a tidygraph object.

Here is a nice tutorial on tidy graph: https://tidygraph.data-imaginist.com/reference/group_graph.html

Just by looking at the network, I can already understand a lot about the position of different actors in a network.

```{r}
library(tidygraph)
library(ggraph)

coll.ch.tidy <- as_tbl_graph(coll.ch.net)

set.seed(123)

coll.ch.tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_authority()) %>% # to display node size based on degree centrality
  ggraph(layout = "graphopt") + 
  geom_edge_link(width = 0.3, colour = "gray") +
  geom_node_point(aes(size = centrality, colour = type)) +
  # geom_node_text(aes(label = klar), repel = TRUE)+
  # scale_color_gradient(low = "yellow", high = "red")+
  theme_graph()
ggsave("Plots/Climate_pol_net.png",  width = 13, height = 10, dpi=300)
```

### What do key actors think about climate policy? 

Previously, we have learnt to investigate basic network structures, to identify the most central, i.e., important or influential actors and we have added some network characteristics to the network.

We have seen, that this allows us to understand, which group of actors, e.g. governmental actors, is more or less central in a network.

We can do the same thing for other characteristics as well! 

Let's try to see what key actors in Switzerland think about _Carbon Pricing_! 

The result is not so clear -- > core actors seem to deviate! 

```{r}
library(tidyverse)

# First calculate the centralization statistics as ususal ...

myCentrality <- function(net) {
  
  if (!is.igraph(net)) stop ("Input is not an igraph object")
  ideg <- degree (net, mode="in", loop=F, normalized = T)
  odeg <- degree (net, mode="out", loop=F, normalized = T)
  btw <- betweenness (net, normalized=TRUE)
  clo <- closeness (net, normalized=TRUE) # consider removing
  evc <- evcent(net)$vector
  id <- V(net)
  
  ret <-data.frame (cbind(IDegree=ideg, ODegree=odeg,
                         Betweenness=btw, Closeness=clo,
                          Eigenvector=evc, ID=id))
  return(ret)
  
}

Results_coll <- myCentrality(coll.ch.net)
Results_coll

# Round
Results_coll_r <- data.frame(apply(Results_coll[1:5],2, function(x) round(x, 3)))
Results_coll_r

# Here we map the another variable, i.e. Carbon Pricing

Results_coll_r$carbonPricing <- beliefs.Ch$Carbon.Pricing

Group_cent_coll <- Results_coll_r %>%
   mutate(carbonPricing = recode(carbonPricing, "1" = "very low support",
         "2" = "low support",
         "3" = "neutral", "4" = "high support", "5"="very high support", "NA" = "none")) %>%
  group_by(carbonPricing) %>%
  summarise(mean_OD = mean(ODegree),  mean_ID = mean(IDegree), mean_Betw = mean(Betweenness), mean_clo=mean(Closeness), mean_Eig=mean(Eigenvector))
Group_cent_coll


```

### Are similar thinking actors connected? - Assortativity

Further, it would be interesting to understand whether actors that think alike (homophily), do also share ties. For this purpose, we can use the `assortativity()` function. Find the documentation here: https://igraph.org/r/doc/assortativity.html

Let's have a look at some belief variables ...

Do actors that share preferences about Regulatory Measures also collaborate more? 

According to the assortativity coefficient: yes

Note: The `assortativity()` function does not accept NAs. This means we need to get rid of them. For the purpose of this tutorial, we will set them to zero. It's, however, debatable, if this is the right approach.

```{r}
beliefs.Ch[is.na(beliefs.Ch)] <- 0

V(coll.ch.net)$Regulatory.Measures <- beliefs.Ch$Regulatory.Measures

Regulations <- as.numeric(factor(V(coll.ch.net)$Regulatory.Measures))

assortativity(coll.ch.net, Regulations)
```
What about subsidies? Yes!

```{r}

V(coll.ch.net)$Subsidies <- beliefs.Ch$Subsidies

Subsidies <- as.numeric(factor(V(coll.ch.net)$Subsidies))

assortativity(coll.ch.net, Subsidies)
```
What about economic competitivness? Yes!

```{r}

V(coll.ch.net)$Comp <- beliefs.Ch$Competitiveness

Comp <- as.numeric(factor(V(coll.ch.net)$Comp))

assortativity(coll.ch.net, Comp)
```

## Exercises No. 3

1. Load the Swiss __influence data__ and create the network.
2. Add the actor type variable from the type.ch dataset, the Effectiveness, Efficiency, Justice, Competitiveness variables form the belief.Ch dataset, and the invol_nego variable from the twolevel.ch dataset.
  2a. What actor types are the most influential ones?
  2b. Do the most influential actors share the same core beliefs?
  2c. Are the most influential actors also invovled in the international climate negotiations?
3.Calculate the assortativity coefficient for three instrument preferences (Regulatory Measures, Technological Innovation, Behavioural Change)
4. Bonus Exercise: Explore a new package to detect __structural holes__. Look at this vignette and play around with the package a bit. For this purpose work with the collaboration network! 

https://search.r-project.org/CRAN/refmans/migraph/html/holes.html


## Solutions No. 3

