---
title: "Script No. 2"
author: "Marlene Kammerer"
date: "12 Oktober 2023"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = TRUE
)

```

# Network centrality {.tabset}

## Tutorial

### Content

  - Load packages 
  - Load the network data
  - Create network objects
  - calculate networks centrality statistics

### Install & load packages

Again, we work with the`igraph` package.

Do not forget to call `library(igraph)`.


```{r}

# install.packages("igraph") # to install the package 
library(igraph) # to load the package

```

### Load the network data

In this tutorial, we work with the US data again. For your exercises, you will work with a brand new Swiss dataset.


For more information on the US data you may have a look at ([Kammerer et al. 2021]https://onlinelibrary.wiley.com/doi/10.1111/psj.12422) 

```{r}
## Attributes

att.usa <- read.csv2("Data/usa_attributes.csv", header=TRUE)

## Collaboration network

coll.usa <- read.csv2("Data/usa_collaboration.csv", header=TRUE)
rownames(coll.usa) <- coll.usa$ID
coll.usa$ID <- NULL

## Influence network

inf.usa <- read.csv2("Data/usa_influence.csv", header = TRUE, stringsAsFactors=FALSE)
rownames(inf.usa) <- inf.usa $ID
inf.usa$ID <- NULL

## Scientific information network

sci.usa <- read.csv2("Data/scientific-information_USA.csv", header = TRUE, stringsAsFactors=FALSE)
rownames(sci.usa) <- sci.usa$ID
sci.usa$ID <- NULL
```

### Explore data sets

As usual, make sure that your datasets are correctly loaded. Check if there are NAs or any other issues that need to be resolved first.

```{r}

View(att.usa)
View(coll.usa)
View(inf.usa)
View(sci.usa)
```

### Create network objects

As usual we have to create network objects! This means to transform the data first into a matrix object (not dataframe) and then into a graph object.

#### First: Create matrices from data frames (only for "networks")

```{r}

coll.mat <- as.matrix(coll.usa)
inf.mat <- as.matrix(inf.usa)
sci.mat <- as.matrix(sci.usa)

# How to deal with NAs?? For the moment, we will recode them to 0. But, at later points we also discuss different strategies for treating NAs.

coll.mat[is.na(coll.mat)] <- 0
inf.mat[is.na(inf.mat)] <- 0
sci.mat[sci.mat == 98] <- 0

```

#### Second: Create network object and visualize it in a first very simple plot

```{r}

coll.mat <- as.matrix(coll.usa)
coll.net <- graph.adjacency(coll.mat, mode="directed", diag=FALSE)
plot(coll.net)

```

### Explore network

Recall, in the last week, we did some basic analyses ...

```{r, eval=TRUE, echo=TRUE}

# number of nodes and edges

vcount(coll.net)
ecount(coll.net)

# cliques

table(sapply(cliques(coll.net), length))

# dyads

dyad.census(coll.net)

# triads

triad.census(coll.net)

# connectivity and paths

is.connected(coll.net)

# components (not really useful here)

comps <- decompose.graph(coll.net)
table(sapply(comps, vcount))

# average path length

average.path.length(coll.net)

diameter(coll.net)

# is this network very connected?

```

### Network centrality

In the following, we calculate different network centrality statistics, such as

- Degree
- Betweeness
- closeness
- Eigenvector centrality 


We use a self-written function to calculate all these stats in one step and save them in a table. Also, we round the results.

```{r, eval=TRUE, echo=TRUE}
myCentrality <- function(net) {
  
  if (!is.igraph(net)) stop ("Input is not an igraph object")
  ideg <- degree (net, mode="in", loop=F, normalized = T)
  odeg <- degree (net, mode="out", loop=F, normalized = T)
  btw <- betweenness (net, normalized=TRUE)
  clo <- closeness (net, normalized=TRUE) # consider removing
  evc <- evcent(net)$vector
  id <- V(net)
  
  ret <-data.frame (cbind(IDegree=ideg, ODegree=odeg,
                         Betweenness=btw, Closeness=clo,
                          Eigenvector=evc, ID=id))
  return(ret)
  
}

Results <- myCentrality(coll.net)
Results

# Round
Results_r <- data.frame(apply(Results[1:5],2, function(x) round(x, 3)))
```


### calculate network stats per actor type

But, what does this all mean? The numbers are for example more informative, if we look at them in an aggregated fashion. So, in this next code-chunk we calculate the average values of these statistics by actor type.

```{r, eval=TRUE, echo=TRUE}
library(tidyverse)

Results_r$actor.type <- as.factor(att.usa$OT_BIG)

Group_cent <- Results_r %>%
   mutate(actor.type = recode(actor.type, "1" = "GOV",
         "2" = "SCI",
         "3" = "BUS", "4" = "CIV", "5"="NA")) %>%
  group_by(actor.type) %>%
  summarise(mean_OD = mean(ODegree),  mean_ID = mean(IDegree), mean_Betw = mean(Betweenness), mean_clo=mean(Closeness), mean_Eig=mean(Eigenvector))
Group_cent

```

## Exercises No. 2

1. Load the Swiss data into R (collaboration network, influence network, beliefs & actor type).
2. Create the collaboration and influence networks!
3. Do basic explorations (like census, diameter, av path length). What do we learn?
4. Assign actor type variable to the networks.
5. Plot the in and out degree distribution for both networks, arrange in one grid. Hint:https://stackoverflow.com/questions/1249548/side-by-side-plots-with-ggplot2
6. Calculate network statistics (e.g. degree, betweenness, closeness, eigenvector) that make sense for your the respective networks - justify your selection. What do the stats mean in the respective context?
7. Which actor group is most central in the collaboration network? Which is perceived as most influential? Who is the most influential actor? Who collaborates most? Who is most the most popular collaboration target?
8. Plot collaboration network with actor type as color code, arcs and degree stats reflecting node sizes. Add a legend to the figure.
9. Save your graph.

## Solutions No. 2

1. Load the Swiss data into R (collaboration network, influence network, beliefs & actor type).

This is a rather easy task, but at the same time, many problems can arise! So always have a look at the datasets your are working with and do not blindly copy any commands. This will usually run into troubles. 

While any programmer will warn you about "line-by-line" coding, I always do that to understand what I am doing and how my data looks like.

```{r}
## Attributes

beliefs.Ch <- read.csv2("Data/CH_beliefs.csv", header=TRUE, fileEncoding = "latin1", stringsAsFactors = FALSE)

## Collaboration network

coll.ch <- read.csv2("Data/cH_coll.csv", header=TRUE)
rownames(coll.ch) <- coll.ch$X
coll.ch$X <- NULL

## Influence network

inf.ch <- read.csv2("Data/cH_influence.csv", header=TRUE)
rownames(inf.ch) <- inf.ch$X
inf.ch$X <- NULL

## Actor type

types.ch <- read.csv2("Data/CH_type.csv", header = TRUE, fileEncoding = "latin1", stringsAsFactors = FALSE)

```

2. Create the collaboration and influence networks!

```{r}
coll.ch.mat <- as.matrix(coll.ch) # transform dataset into matrix format
coll.ch.net <- graph.adjacency(coll.ch.mat, mode="directed", diag=FALSE) # transform matrix into graph object

inf.ch.mat <- as.matrix(inf.ch) # transform dataset into matrix format
inf.ch.net <- graph.adjacency(inf.ch.mat, mode="directed", diag=FALSE) # transform matrix into graph object
```

3. Do basic explorations (like census, diameter, av path length). What do we learn?

I simply run all the analyses from the tutorial and I do this now only for the collaboration network. 

I can see that the collaboration network has 54 nodes, and 465 edges. It has 84 mutual connections, but 279 asymmetric connections, i.e. directed ties. Also, it has 1050 potential ties that are NOT present. Also, the network is composed of one single component. The diameter is 5 (longest connection) and an average path length of 1.94. Think about the diameter: Have you heard about the small world phenomenon? See here: https://en.wikipedia.org/wiki/Small-world_experiment. According this this we are all only six degrees (on average) away from each other. Our network is much more connected, but looking at the diameter (the furthest connection), this network is pretty close.

```{r, eval=TRUE, echo=TRUE}

# number of nodes and edges

vcount(coll.ch.net)
ecount(coll.ch.net)

# cliques

table(sapply(cliques(coll.ch.net), length))

# dyads

dyad.census(coll.ch.net)

# triads

triad.census(coll.ch.net)

# connectivity and paths

is.connected(coll.ch.net)

# components (not really useful here)

comps <- decompose.graph(coll.ch.net)
table(sapply(comps, vcount))

# average path length

average.path.length(coll.ch.net)

diameter(coll.ch.net)

```
4. Assign actor type variable to the networks.

```{r, eval=TRUE, echo=TRUE}

V(coll.ch.net)$type <- types.ch$Type
V(coll.ch.net)$klar <- types.ch$Klarname_EN
list.vertex.attributes(coll.ch.net)

V(inf.ch.net)$type <- types.ch$Type
V(inf.ch.net)$klar <- types.ch$Klarname_EN
list.vertex.attributes(inf.ch.net)
```
5. Plot the in and out degree distribution for both networks, arrange in one grid. Hint:https://stackoverflow.com/questions/1249548/side-by-side-plots-with-ggplot2

First for the collaboration network ...

We see that most actors few to a medium number of incoming ties. Hence, very few actor are very popular or very unpopular collaboration targets.
Further, we see that most actors have very few or few outgoing ties. Hence, very few actors are very active collaborators.


```{r, eval=TRUE, echo=TRUE}

par(mfrow=c(1,2))

hist (degree(coll.ch.net, mode = "in", normalized = TRUE), col="yellow", xlim=c(0,0.4),
      xlab = "Vertex Indegree", ylab="Frequency", main= "InDegree Distribution Switzerland")

hist (degree(coll.ch.net, mode = "out", normalized = TRUE), col="yellow", xlim=c(0,0.8),
      xlab = "Vertex Outdegree", ylab="Frequency", main= "OutDegree Distribution Switzerland")
```

Now for the influence network.

What do we see? 

We see that the figures look very similar. Most actors are not percieved as influential, vice versa, and what makes sense, only few actors are considered as being influential.

```{r, eval=TRUE, echo=TRUE}

par(mfrow=c(1,2))

hist (degree(inf.ch.net, mode = "in", normalized = TRUE), col="blue", xlim=c(0,0.8),
      xlab = "Vertex Indegree", ylab="Frequency", main= "InDegree Distribution Switzerland")

hist (degree(inf.ch.net, mode = "out", normalized = TRUE), col="blue", xlim=c(0,0.8),
      xlab = "Vertex Outdegree", ylab="Frequency", main= "OutDegree Distribution Switzerland")
```

6. Calculate network statistics (e.g. degree, betweenness, closeness, eigenvector) that make sense for your the respective networks - justify your selection. What do the stats mean in the respective context?

Again, we can replicate the code! I run them all for the collaboration network...

I have NaNs in the Closeness Centrality statistic. Why is that?

Which centrality stats make sense for the influence network?

```{r, eval=TRUE, echo=TRUE}
myCentrality <- function(net) {
  
  if (!is.igraph(net)) stop ("Input is not an igraph object")
  ideg <- degree (net, mode="in", loop=F, normalized = T)
  odeg <- degree (net, mode="out", loop=F, normalized = T)
  btw <- betweenness (net, normalized=TRUE)
  clo <- closeness (net, normalized=TRUE) # consider removing
  evc <- evcent(net)$vector
  id <- V(net)
  
  ret <-data.frame (cbind(IDegree=ideg, ODegree=odeg,
                         Betweenness=btw, Closeness=clo,
                          Eigenvector=evc, ID=id))
  return(ret)
  
}

Results_coll <- myCentrality(coll.ch.net)
Results_coll

# Round
Results_coll_r <- data.frame(apply(Results_coll[1:5],2, function(x) round(x, 3)))
Results_coll_r
```
```{r, eval=TRUE, echo=TRUE}

Results_inf <- myCentrality(inf.ch.net)
Results_inf

# Round
Results_inf_r <- data.frame(apply(Results_inf[1:5],2, function(x) round(x, 3)))
Results_inf_r
```
7. Which actor group is most central in the collaboration network? Which is perceived as most influential? Who is the most influential actor? Who collaborates most? Who is most the most popular collaboration target?

```{r, eval=TRUE, echo=TRUE}
library(tidyverse)

Results_coll_r$actor.type <- types.ch$Type

Group_cent_coll <- Results_r %>%
   mutate(actor.type = recode(actor.type, "1" = "GOV",
         "2" = "SCI",
         "3" = "BUS", "4" = "CIV", "5"="NA")) %>%
  group_by(actor.type) %>%
  summarise(mean_OD = mean(ODegree),  mean_ID = mean(IDegree), mean_Betw = mean(Betweenness), mean_clo=mean(Closeness), mean_Eig=mean(Eigenvector))
Group_cent_coll
```

```{r, eval=TRUE, echo=TRUE}
library(tidyverse)

Results_inf_r$actor.type <- types.ch$Type

Group_cent_inf <- Results_r %>%
   mutate(actor.type = recode(actor.type, "1" = "GOV",
         "2" = "SCI",
         "3" = "BUS", "4" = "CIV", "5"="NA")) %>%
  group_by(actor.type) %>%
  summarise(mean_OD = mean(ODegree),  mean_ID = mean(IDegree), mean_Betw = mean(Betweenness), mean_clo=mean(Closeness), mean_Eig=mean(Eigenvector))
Group_cent_inf
```

```{r, eval=TRUE, echo=TRUE}
detach(package:igraph) # igraph will cause confusion with other packages, hence we need to detach it.
library(sna) # another network package; 
library(GGally)
library(network) # another network package; 
library(ggplot2) # for plotting
library(RColorBrewer)
library(intergraph) # another network package; 
library(scales) 
library(tidyverse) # data management

net <- network(coll.ch.mat, directed = TRUE) # this is a different way to create a graph object (using the network package);

# To add attributes here:

net%v%"actorType" <- types.ch$Type

unique(types.ch$Type)

# Adjust your color palette (tip: https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3)!
# Unique actor types: "Business" "Civil society" "Government"    "Science"       "Party"     

# Exmple on how to create a color palette for the actor type variable

palette = c("Government" = "#f03b20", "Science" = "#31a354", 
             "Business" ="#feb24c", "Civil society" = "#c51b8a", "Party" = "#2b8cbe")

# Adjust this code to your taste! If you don't know what to do, go and ask Dr. Google!

ggnet2(net, alpha = 0.75, label = FALSE, label.size = 3, mode = "kamadakawai", node.size = 6, size = "indegree", color = "actorType", 
       arrow.size = 8, arrow.gap = 0.025,
       shape.legend = "Actor Type",
       color.legend = "actorType", edge.color = "grey", # maybe you have to adjust the name of the actorType variable here
       palette = palette, 
       legend.size = 10, size.min = 1, size.cut = 5, size.legend = "8", legend.position = "bottom")
ggsave("Plots/my_CH_network.png", width = 29, height = 18,  device = NULL, dpi = 300)
```